---
title: "BIOL 432 Assignment 1 R Markdown"
author: "Laavanya Joshi"
date: "Monday, January 6, 2024"
output:
  html_document:
    df_print: paged
---

## [Assignment1Github](https://github.com/LaavanyaJoshi/BIOL432-Assignment-1)


### Executes RScripts 1 and 2, respectively, which opens the original and updated datasets of Butterfly species
```{r}
source("C:/Users/laava/OneDrive/Documents/Queen's University/Fourth Year/BIOL 432/BIOL432-Assignment-1/dataGenerato.R")
source("C:/Users/laava/OneDrive/Documents/Queen's University/Fourth Year/BIOL 432/BIOL432-Assignment-1/volumeEstimato.R")
```


### Installs the required packages and uploads the csv datafile this analysis will use
```{r}
library(dplyr)
library(ggplot2)

measurements <- read.csv("C:/Users/laava/OneDrive/Documents/Queen's University/Fourth Year/BIOL 432/BIOL432-Assignment-1/measurements.csv")
```


### Sorts the data by Species, then Observer, then Limb Volume
```{r}
sorted_data <- measurements %>%
  arrange(Species, Observer, Limb_volume)
```


### View a table of Average Limb Volume for each Species
```{r}
measurements %>% 
  count(Species, mean(Limb_volume)) %>%
  setNames(c("Species", "Average Limb Volume (mm^3)", "Count"))
```


### View a table of number of Species observed by each person
```{r}
measurements %>%
  count(Species, Observer) %>%
  setNames(c("Species", "Observer", "Count"))
```


### Creates a boxplot showing the distribution of limb volume for each species
```{r}
ggplot(data = measurements) + 
  geom_boxplot(aes(x = Species, y = Limb_volume)) + 
  theme_bw() + 
  labs(x = "Species", y = "Limb Volume (mm^3)", title = "Boxplot of Distribution of Limb Volume Measurements for Butterfly Observations")
```


### Creates a faceted histogram to compare limb volumes for each species
```{r}
ggplot(data = measurements) + 
  geom_histogram(aes(x = Limb_volume), bins = 25) + 
  facet_wrap(~Species, nrow = 3, ncol = 2, scales = "free_x") + 
  theme_bw() + 
  labs(x = "Limb Volume (mm^3)", y = "Counts", title = "Composite Plot of Histograms showing Limb Volumes for Butterfly Observations")
```

